---
title: '**R** codes for the case study of the paper *Computing and visualizing impacts
  in the general spatial interaction model*'
author: "Thibault Laurent & Paula Margaretic & Christine Thomas-Agnan"
date: 'Last update: `r Sys.Date()`'
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 2
    keep_tex: yes
---

```{r setup, include=FALSE}

library(knitr)
# ORIGINAL: knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(echo = TRUE,dev = c('pdf', 'png'), 
        fig.align = 'center', fig.height = 5, fig.width = 8.5)

#Determine the output format of the document
outputFormat   = opts_knit$get("rmarkdown.pandoc.to")

#Figure and Table Caption Numbering, for HTML do it manually
capTabNo = 1; capFigNo = 1;

#Function to add the Table Number
capTab = function(x){
  if(outputFormat == 'html'){
    x = paste0("Table ",capTabNo,". ",x)
    capTabNo <<- capTabNo + 1
  }; x
}

#Function to add the Figure Number
capFig = function(x){
  if(outputFormat == 'html'){
    x = paste0("Figure ",capFigNo,". ",x)
    capFigNo <<- capFigNo + 1
  }; x
}
```


This document provides the **R** codes used to reproduce the new case study of the paper  *Computing and visualizing impacts in the general spatialinteraction model*. The new case study only considers European countries. The pdf version is available [here](index.pdf).

The dataset is rectangular. It comprises 17 European countries: 13 of them act both as origins and destinations, 2 countries are only origins and 2 are only destinations.
The countries are the following: Austria, Bulgaria,    Switzerland, Czech Republic, Germany, Spain, France, United Kingdom,   Hungary, Ireland, Italy, Netherlands, Norway,          Poland,Portugal, Romania,  Russia.  

Packages needed:
```{r, message = F, eval = F,echo=FALSE}
install.packages(c(
  "dplyr", # modify, select data
  "gridExtra", # several ggplot
  "ggalluvial", # spatial flows
  "haven", # import stat files
  "igraph", # library for network
  "kableExtra", # table with Markdown
  "lwgeom", "sf", "spdep", # spatial packages
  "Matrix", # sparse matrix
  "plyr", # modify object
  "viridis", # data vizualisation
  "hrbrthemes" # data vizualization
  )) 
devtools::install_github('rensa/ggflags')
```


```{r, message = F,echo=FALSE}
library("viridis")
library("hrbrthemes")
library("gridExtra")
library("ggflags")
library("ggalluvial")
library("kableExtra")
library("tidyverse")
library("sf")
library("spdep")
library("Matrix")  
```

# Treatment of the data

## Import remittances data

We import the remittances data:

```{r, echo = F, eval = F}
# my_dta <- haven::read_dta("./data/Remittances_v20200110.dta")

my_dta <- haven::read_dta("./data/Remittances_v20200110.dta")

# We subset the years of interest (2012 and 2017) and select the variables of interest:

my_dta_1 <- my_dta %>%
  filter(year == 2012 | year == 2017) %>%
  select(1:4, 8, 12, 13, 15, 17:20, 27:31)

# We set to 0 the NA values on remittances (it concerns 4 countries, in 2012):

my_dta_1[which(is.na(my_dta_1$remit)), "remit"] <- 0

# We identify countries that have not received money in 2012 or 2017:

id_norec_2012_17 <- sort(unique(c(
  my_dta_1 %>% 
    filter(year == 2012) %>%
    group_by(reccode2) %>% 
    dplyr::summarise(remit_cum = sum(remit)) %>% 
    filter(remit_cum == 0) %>% 
    select(reccode2) %>%
    pull(reccode2),
  my_dta_1 %>% 
    filter(year == 2017) %>%
    group_by(reccode2) %>% 
    dplyr::summarise(remit_cum = sum(remit)) %>% 
    filter(remit_cum == 0) %>% 
    select(reccode2) %>%
    pull(reccode2)
)))


# We identify countries that have not sent money in 2012 or 2017:

id_nosend_2012_17 <- sort(unique(c(
  my_dta_1 %>% 
    filter(year == 2012) %>%
    group_by(sendcode2) %>% 
    dplyr::summarise(remit_cum = sum(remit)) %>% 
    filter(remit_cum == 0) %>% 
    select(sendcode2) %>%
    pull(sendcode2),
  my_dta_1 %>% 
    filter(year == 2017) %>%
    group_by(sendcode2) %>% 
    dplyr::summarise(remit_cum = sum(remit)) %>% 
    filter(remit_cum == 0) %>% 
    select(sendcode2) %>%
    pull(sendcode2)
)))


# We create the list of European countries we will use in this case study
codes<-unique(my_dta_1[,c(1,4,7)])

list_O<-c("AT","CZ","FR","DE","HU","IE","IT","NL","NO","PT","RO","RU","ES","CH","GB")
list_D<-c("AT","BG","CZ","FR","DE","HU","IT","NL","PL","PT","RO","RU","ES","CH","GB")

my_dta_2 <- my_dta_1 %>%  
  filter(reccode2 %in% list_D,
         sendcode2 %in% list_O)


## Explanatory variables 

# We import the explanatory variables: 

varexp <- haven::read_dta("./data/varexplicativas_v20201029.dta")
varexp1 <- varexp %>%
  filter(year == 2010 | year == 2015) %>%
  select(1, 3, 39, 2, 4:9, 14:16, 18:29, 35, 40, 90:91, 93:94,
        95:98, 107, 110, 111, 112:115, 117, 118, 121, 123:125, 128)

# We give name variables to the last variables chosen:
colnames(varexp1)[c(43:49)]<-c("new_bus", "new_bus_1000", "unemployment_percent",
                               "GDP_growth", "HDI", "HDI_wpcInc", "compensat_USD")
varexp1[,c(43:46)] <- sapply(varexp1[,c(43:46)], as.numeric)
summary(varexp1)

# We Merge electricity use data with varexp1

elect_2016 <- subset(varexp, varexp$year == 2016, select = c(3, 126:127))
varexp1 <- merge(varexp1, elect_2016, by.x = "CountryCode",
                      by.y = "CountryCode", all.x = T, all.y = T)
colnames(varexp1)[50:51] <- c("Electricity_use", "Electricity_use_pc")


# We keep the info of the European countries
id_countries<-union(list_D,list_O)
id_countries_3<-data.frame(codes[codes$sendcode2 %in% id_countries,1])$countrycode1
varexp2 <- varexp1[varexp1$CountryCode %in% id_countries_3, ]
varexp2 <- varexp2[ , -c(11:13)]

# We instrument the transfer cost (as proxied by the fee of a 200 USD transfer made with Western Union, hereafter WU) using the financial development index of the country (variable FD) and whether the country is under a dollarization regime (variable dollarization):

dol_FD_2013 <- subset(varexp, varexp$year == 2013, select = c(3, 39, 40, 97))
dol_FD_2013_1 <- dol_FD_2013[dol_FD_2013$CountryCode %in% id_countries_3, ]
WU_2015 <- subset(varexp, varexp$year == 2015, select = c(3, 39, 4))
WU_2015_1 <- WU_2015[WU_2015$CountryCode %in% id_countries_3, ]
data_WU_inst <- merge(WU_2015_1, dol_FD_2013_1, by.x = "CountryCode",
                      by.y = "CountryCode", all.x = T, all.y = T)
reg_WU_inst <- lm(westunion ~ FD + dollarization, data = data_WU_inst)
data_WU_inst$WU_pred <- predict(reg_WU_inst)
data_WU_inst$WU_res <- residuals(reg_WU_inst)

## Spatial data

# We import the contours of the countries:

world <- sf::st_read("./data/World WGS84/Pays_WGS84.shp")

# We have to add Hong-Kong in the spatial data (as there was an issue in the data). 

library("sf")

## Focus on 2017

# We subset year 2017. We first merge the "remittances" and the "explanatory variables" datasets for this year. The explanatory variables are lagged two years, thus corresponding to year 2015:


my_dta_5 <- merge(my_dta_2[which(my_dta_2$year == "2017"), ],
                  varexp2[which(varexp2$year == "2015"), ],
                  by.x = "countrycode1", by.y = "CountryCode",
                  all.x = T)

# We change the names of some explanatory variables:

names_var <- c("deflactor", "WU", "controls", "dualFX", "multFX", "restrictions", 
               "migration",  "ocurEarthq", "totafEarthq", "totdamEarthq", "dumEarthq",
               "ocurFlood", "totafFlood", "totdamFlood", "dumFlood", "ocurStorm", "totafStorm", 
               "totdamStorm", "dumStorm",     "GDP_PPP", "FD", 
               "island", "landlocked", "GDP", "population", "GDPconstant", "region", "dollarization", 
               "offdevelopment", "polstab", "FXavg","FXsd", "GDP_percapita_UN", "GDP_UN", "GDPconstant_UN",
               "deflactor_UN", "new_bus" ,"new_bus_1000" , "unemployment_percent", "GDP_growth",
               "HDI", "HDI_wpcInc", "compensat_USD", "Electricity_use", "Electricity_use_pc")
colnames(my_dta_5)[20:64] <- paste0("send_", names_var) 

# We add the WU variable which was created independently:

data5_bis <- merge(my_dta_5, data_WU_inst, by.x = "countrycode1", 
                   by.y = "CountryCode", all.x = T)
data5_bis <- data5_bis[ , -c(65:69)]
colnames(data5_bis)[c(65:66)] <- c("send_WU_pred", "send_WU_resid")

# We now add the control variables for the countries at origin and change the names:

data5_bis2 <- merge(data5_bis, varexp2[which(varexp2$year == "2015"),], 
                    by.x = "countrycode2", by.y = "CountryCode", all.x = T)
colnames(data5_bis2)[c(69:113)] <- paste0("rec_", names_var) 
data5_bis2 <- data5_bis2[, -c(18, 19, 67, 68)]

# We first identify and eliminate those countries which have too few remittances. Specifically, we consider the year 2017 and search which countries have zero remittances with at least 110 counterparts. We then eliminate them. Also, we eliminate the zero  (off-diagonal) flows.

my_dta_6 <- data5_bis2
my_dta_6 <- my_dta_6[order(my_dta_6$send_ct, my_dta_6$rec_ct),]

# my_dta_6$zero_fl <- F
# my_dta_6$zero_fl[which(my_dta_6$remit == 0)] <- T
# 
# id_orig_remit_zero <- my_dta_6 %>% 
#   filter(zero_fl) %>%
#   group_by(sendcode2) %>% 
#   summarise(zero_cum = n()) %>%
#   filter(zero_cum > 110) %>%
#   select(sendcode2) %>%
#   pull(sendcode2)
# 
# id_dest_remit_zero <- my_dta_6 %>% 
#   filter(zero_fl) %>%
#   group_by(reccode2) %>% 
#   summarise(zero_cum = n()) %>%
#   filter(zero_cum > 110) %>%
#   select(reccode2) %>%
#   pull(reccode2)
# 
# # Eliminate those countries without enough remittances data 
# 
# my_dta_6 <- my_dta_6[!my_dta_6$sendcode2 %in% id_orig_remit_zero,]
# my_dta_6 <- my_dta_6[!my_dta_6$reccode2 %in% id_dest_remit_zero,]
# 

## Computing the distances 

# The distances between flows can be presented in a matrix of size $n_d \times n_o$. We use the function *gDistance()* from package **rgeos** (Bivand and Rundel, 2019).

id_send_ct <- sort(unique(my_dta_6$send_ct))
id_rec_ct <- sort(unique(my_dta_6$rec_ct))

# Create a dataframe with the country names and their ISO country codes (2 digits). Add the ISO codes to the geometry object to be able later to sort

id_send_ct_nam_iso <- unique(data.frame(my_dta_6$send_ct,my_dta_6$sendcode2))
colnames(id_send_ct_nam_iso) <- c("send_ct", "sendcode2")
id_send_ct_nam_iso <- id_send_ct_nam_iso[order(id_send_ct_nam_iso$sendcode2),]

id_rec_ct_nam_iso <- unique(data.frame(my_dta_6$rec_ct, my_dta_6$reccode2))
colnames(id_rec_ct_nam_iso) <- c("rec_ct", "reccode2")
id_rec_ct_nam_iso <- id_rec_ct_nam_iso[order(id_rec_ct_nam_iso$reccode2),]

# Create the geometry object and add the ISO codes 

geom_send_ct0 <- world[world$NOM %in% id_send_ct, ]
geom_send_ct <- merge(geom_send_ct0, id_send_ct_nam_iso, 
                      by.x = "NOM", by.y = "send_ct", all.x = TRUE)
row.names(geom_send_ct) <- geom_send_ct$sendcode2
geom_send_ct <- geom_send_ct[order(geom_send_ct$sendcode2), ]

geom_rec_ct0 <- world[world$NOM %in% id_rec_ct, ]
geom_rec_ct <- merge(geom_rec_ct0, id_rec_ct_nam_iso, 
                     by.x = "NOM", by.y = "rec_ct", all.x = TRUE)
row.names(geom_rec_ct) <- geom_rec_ct$reccode2
geom_rec_ct <- geom_rec_ct[order(geom_rec_ct$reccode2), ]


library(lwgeom)
G_mig <- t(as(st_distance(st_centroid(geom_send_ct), 
                          st_centroid(geom_rec_ct)), "matrix"))

# It can also be  added to the data.frame which presents the data in vectorized form

row.names(G_mig) <- row.names(geom_rec_ct)
colnames(G_mig) <- row.names(geom_send_ct)


## Construct the spatial weight matrices 

# To define the spatial weight matrix on our geographical sites, we use the 3-nearest neighbours properties for origins and destinations. To store the $R$ explanatory variables in a matrix of size $N\times R$, the user has to use Kronecker products separately for origins and for destinations: 

n_rem_o <- length(unique(my_dta_6$sendcode2))
n_rem_d <- length(unique(my_dta_6$reccode2))
n_rem_o * n_rem_d == nrow(my_dta_6)

### Matrix Wo

# First, we build the spatial weight matrix on the countries which are observed at the origin. The matrix is based on the 3-nearest neighbors. 

library("spdep")
w_rem_o_nb <- knn2nb(knearneigh(as(st_centroid(geom_send_ct), "Spatial"), k = 3))
w_rem_o <- listw2mat(nb2listw(w_rem_o_nb))

# We also compute a matrix of distances between observations at origin (we will use it later to correct some neighbours).

w_rem_o_dist <- t(as(st_distance(st_centroid(geom_send_ct), 
                                 st_centroid(geom_send_ct)), "matrix"))
row.names(w_rem_o_dist) <- row.names(w_rem_o) <- row.names(geom_send_ct)
colnames(w_rem_o_dist) <- colnames(w_rem_o) <- row.names(geom_send_ct)

# To build the spatial weight matrices $W_o$, we use Kronecker products. 

library("Matrix")
W_rem_o <- kronecker(w_rem_o, Diagonal(n_rem_d))
W_rem_o_dist <- kronecker(w_rem_o_dist, Diagonal(n_rem_d))
W_rem_o_list <- mat2listw(W_rem_o, style = "W")
row.names(W_rem_o) <- row.names(W_rem_o_dist) <- 
  paste(rep(id_send_ct_nam_iso$sendcode2, each = n_rem_d), 
        rep(id_rec_ct_nam_iso$reccode2, n_rem_o), sep = " ")
colnames(W_rem_o) <-colnames(W_rem_o_dist) <- 
  paste(rep(id_send_ct_nam_iso$sendcode2, each = n_rem_d), 
        rep(id_rec_ct_nam_iso$reccode2, n_rem_o), sep = " ")

# We take the neighbours (still a list)
lwo_neighb <- W_rem_o_list$neighbours


### Matrix Wd

# We build the spatial weight matrix on the countries which are observed at the destination. The matrix is based on the 3-nearest neighbors. 

w_rem_d_nb <- knn2nb(knearneigh(as(st_centroid(geom_rec_ct), "Spatial"), k = 3))
w_rem_d <- listw2mat(nb2listw(w_rem_d_nb))

# We compute the distance matrix:

w_rem_d_dist <- t(as(st_distance(st_centroid(geom_rec_ct),
                                 st_centroid(geom_rec_ct)), "matrix"))
row.names(w_rem_d_dist) <- row.names(w_rem_d) <- row.names(geom_rec_ct)
colnames(w_rem_d_dist) <- colnames(w_rem_d) <- row.names(geom_rec_ct)

# Finally, we use Kronecker product to build $W_d$:

W_rem_d <- kronecker(Diagonal(n_rem_o), w_rem_d)
W_rem_d_dist <- kronecker(Diagonal(n_rem_o), w_rem_d_dist)
W_rem_d_list <- mat2listw(W_rem_d, style = "W")
row.names(W_rem_d) <- row.names(W_rem_d_dist) <- 
  paste(rep(id_send_ct_nam_iso$sendcode2, each = n_rem_d), 
        rep(id_rec_ct_nam_iso$reccode2, n_rem_o), 
        sep = " ")
colnames(W_rem_d) <-colnames(W_rem_d_dist) <- 
  paste(rep(id_send_ct_nam_iso$sendcode2, each=n_rem_d), 
        rep(id_rec_ct_nam_iso$reccode2, n_rem_o), sep=" ")

# We take the neighbours (still a list)
lwd_neighb <- W_rem_d_list$neighbours

### Matrix Ww

# We build $W_w$ thanks to $W_o$ and $W_d$

W_rem_w <- W_rem_o %*% W_rem_d
W_rem_w_list <- mat2listw(W_rem_w, style = "W")
lww_neighb <- W_rem_w_list$neighbours


## Transform the explanatory variables to origin-destination format 

# We order the data basis:

my_dta_6 <- my_dta_6[order(my_dta_6$sendcode2, my_dta_6$reccode2),]
rownames(my_dta_6) <- seq(1:nrow(my_dta_6))
my_dta_6$rec_GDP_PPP <- as.numeric(as.character(my_dta_6$rec_GDP_PPP))
my_dta_6$send_GDP_PPP <- as.numeric(as.character(my_dta_6$send_GDP_PPP))

# We create some additional variables which are dummies:

my_dta_6$rec_dum_nat_dis <- 0
my_dta_6$rec_dum_nat_dis[which(my_dta_6$rec_dumEarthq == 1 | 
                                 my_dta_6$rec_dumStorm == 1)] <- 1
my_dta_6$send_cont_mulFX <- 0
my_dta_6$send_cont_mulFX[which(my_dta_6$send_controls == 1 | 
                                 my_dta_6$send_dualFX == 1 | 
                                 my_dta_6$send_multFX == 1)] <- 1
my_dta_6$rec_cont_mulFX <- 0
my_dta_6$rec_cont_mulFX[which(my_dta_6$rec_controls == 1 | 
                                my_dta_6$rec_dualFX == 1 | 
                                my_dta_6$rec_multFX == 1)] <- 1
my_dta_6$same_region <- 0
my_dta_6$same_region[which(my_dta_6$rec_region == my_dta_6$send_region)] <- 1

# We separate the explanatory variables at Origin:
X_o <- as.matrix(my_dta_6 %>% 
                   select(send_GDPconstant_UN, send_population, send_FD, send_deflactor_UN,
                          send_island, send_landlocked, contig,comlang_off, comcur, colony,
                          same_region, send_controls, send_dualFX, send_multFX, send_cont_mulFX,
                          send_migration,  send_FXavg, send_FXsd,
                          send_polstab, send_dollarization, send_WU_pred, send_WU_resid,
                          send_new_bus_1000, send_new_bus,
                          send_GDP_growth, send_HDI_wpcInc, send_unemployment_percent,
                          send_Electricity_use, send_Electricity_use_pc))
rownames(X_o) <- my_dta_6$sendcode2

# We separate the explanatory variables at Destination:

X_d <- as.matrix(my_dta_6 %>% 
                   dplyr::select(rec_GDPconstant_UN, rec_population, rec_FD, rec_deflactor_UN, rec_island, 
                                 rec_landlocked, rec_controls, rec_dualFX, rec_multFX, rec_cont_mulFX, 
                                 rec_migration, rec_FXavg, rec_FXsd, rec_ocurEarthq,
                                 rec_totafEarthq, rec_totdamEarthq, rec_dumEarthq, 
                                 rec_ocurFlood, rec_totafFlood, rec_totdamFlood, rec_dumFlood, rec_ocurStorm,
                                 rec_totafStorm, rec_totdamStorm, rec_dumStorm,
                                 rec_dum_nat_dis, rec_polstab, rec_dollarization,
                                 rec_new_bus_1000, rec_new_bus,
                                 rec_GDP_growth, rec_HDI_wpcInc, rec_unemployment_percent,
                                 rec_Electricity_use, rec_Electricity_use_pc))

# We include the $X_o$ and $X_d$ with the names of origin and destination in the final data basis:

flows_rem <- data.frame(origin = rep(unique(my_dta_6$sendcode2), each = n_rem_d),
                        dest = rep(unique(my_dta_6$reccode2), n_rem_o),
                        X_o, X_d)

# We convert into log some variables and change the names:

flows_rem_log <- as.data.frame(lapply(flows_rem[, c(4, 25:26, 30, 33, 60:61, 65)], 
                                      function(x) log(x)))
colnames(flows_rem_log) <- c("log_sendpop", "log_send_new_bus_1000", "log_send_new_bus", "log_send_elect_use",
                             "log_rec_pop", "log_rec_new_bus_1000",
                             "log_rec_new_bus", "log_rec_Electricity_use")
flows_rem <- cbind(flows_rem, flows_rem_log)

# We transform the matrix of distances in a vector and we add it in the data:

flows_rem$g <- as.vector(G_mig)

# We add the dependent variable and create the logarithms:

flows_rem$y <- my_dta_6$remit
flows_rem$logy <- log(flows_rem$y)
flows_rem$logg <- log(flows_rem$g)

# We add two explanatory variables to the dataset:
flows_rem$log_send_gdp_pop <- log(flows_rem$send_GDPconstant_UN / 
                                    flows_rem$send_population)
flows_rem$log_rec_gdp_pop <- log(flows_rem$rec_GDPconstant_UN / 
                                   flows_rem$rec_population)

# We give an unique name for each flow:
flows_rem$ct_p <- paste(flows_rem$origin, flows_rem$dest, sep = " ")
row.names(flows_rem) <- flows_rem$ct_p


## Eliminate the zero flows 

# In short, we identify the zero flows and we then eliminate them

# 1) Identify the row numbers of the zero  flows 
id_zero_fl <- which(flows_rem$y == 0)
flows_rem.red <- flows_rem[-id_zero_fl, ]

# 2) Procedure for elimination of the zero flows 
# 
# * Eliminate them from the **listw**. Do the subset

lwo_neighb_red <- subset(lwo_neighb, !(1:length(lwo_neighb) %in% id_zero_fl))
attr(lwo_neighb_red, "country.id") <- seq(1:length(lwo_neighb_red))

lwd_neighb_red <- subset(lwd_neighb, !(1:length(lwd_neighb) %in% id_zero_fl))
attr(lwo_neighb_red, "country.id") <- seq(1:length(lwo_neighb_red))

lww_neighb_red <- subset(lww_neighb, !(1:length(lww_neighb) %in% id_zero_fl))
attr(lww_neighb_red, "country.id") <- seq(1:length(lww_neighb_red))


# We convert the **nb** list to matrix,  convert the **matrix** to  **list** and eliminate the matrix. Problem : some flows do not have neighbours.
# 
# * For $W_o$

LWo_3_red <- nb2mat(lwo_neighb_red, glist = NULL, style = "W", zero.policy = TRUE)
lwo_3_red <- mat2listw(LWo_3_red, style = "W")
LWo_3_red <- NULL

# We use a sparse **Matrix** object:

W_o_3_red <- as(as_dgRMatrix_listw(lwo_3_red), "CsparseMatrix")
row.names(W_o_3_red) <- colnames(W_o_3_red) <- row.names(flows_rem.red)

# We also drop the 0 from the distance matrix:

W_o_3_dist_red <- W_rem_o_dist[-id_zero_fl, -id_zero_fl]
row.names(W_o_3_dist_red) <- colnames(W_o_3_dist_red) <- row.names(flows_rem.red)


# * We do the same thing for $W_d$:

LWd_3_red <- nb2mat(lwd_neighb_red, glist = NULL, style = "W", zero.policy = TRUE)
lwd_3_red <- mat2listw(LWd_3_red,style = "W")
LWd_3_red <- NULL
W_d_3_red <- as(as_dgRMatrix_listw(lwd_3_red), "CsparseMatrix")
row.names(W_d_3_red) <- colnames(W_d_3_red) <- row.names(flows_rem.red)
W_d_3_dist_red <- W_rem_d_dist[-id_zero_fl, -id_zero_fl]
row.names(W_d_3_dist_red) <- colnames(W_d_3_dist_red) <- row.names(flows_rem.red)

# * We do the same thing for $W_w$:

LWw_3_red <- nb2mat(lww_neighb_red, glist = NULL, style = "W", zero.policy = TRUE)
lww_3_red <- mat2listw(LWw_3_red,style="W")
LWw_3_red <- NULL

##  Deal with the zero flows and those without neighbours after the subsetting 

# We use code "Zero_flows_W_red_threshold_neut.R" to get the new weight matrices **W** (including distance thresholds and subset of flows with no nb)

source("./codes/Zero_flows_W_red_threshold_neut.R")


# Lagged explanatory variables in 2017

Wo.dgT_o <- Wo.dgT_o / apply(Wo.dgT_o, 1, sum)
Wd.dgT_d <- Wd.dgT_d / apply(Wd.dgT_d, 1, sum)


# We create dummy variables for high, medium and low inflation regimes, with the thresholds based on the 10 and 90% percentiles of the distribution 

# * Inflation regime for sending countries. In the European case, we consider high or low
quantile(flows_rem.red$send_deflactor_UN, probs = c(0.10, 0.5,0.75, 0.9))
flows_rem.red$send_high_inf <- 0
flows_rem.red$send_high_inf[which(flows_rem.red$send_deflactor_UN >= 2.46)] <- 1

# flows_rem.red$send_med_inf <- 0
# flows_rem.red$send_med_inf[which(flows_rem.red$send_deflactor_UN < 7.6 & flows_rem.red$send_deflactor_UN >= -0.6)] <- 1

# * Inflation regime for receiving countries

quantile(flows_rem.red$rec_deflactor_UN, probs = c(0.10, 0.5,0.75, 0.9))
flows_rem.red$rec_high_inf <- 0
flows_rem.red$rec_high_inf[which(flows_rem.red$rec_deflactor_UN >= 2.45)] <- 1

# flows_rem.red$rec_med_inf <- 0
# flows_rem.red$rec_med_inf[which(flows_rem.red$rec_deflactor_UN < 2.6 & flows_rem.red$rec_deflactor_UN >= 0.6)] <- 1


save(flows_rem.red, geom_send_ct, geom_rec_ct, flow.no.neigh.o, 
     flow.no.neigh.d, Wd.dgT_d, Wo.dgT_o, file = "my_data.RData")
```

```{r}
# Load the data again
 load(file = "my_data.RData")
```

# Data management

We add some variables and we create the geometries of the unique countries:
```{r,echo=FALSE,message=FALSE,warning=FALSE}
geom_send_ct <- geom_send_ct %>%
  mutate(code = sendcode2)
geom_rec_ct <- geom_rec_ct %>%
  mutate(code = reccode2)
sf_rem_o_d <- rbind(
  geom_send_ct[, c("NOM", "code")],
  geom_rec_ct[!geom_rec_ct$NOM %in% geom_send_ct$NOM, c("NOM", "code")])


```




```{r,echo=FALSE,message=FALSE,warning=FALSE}
# We drop the flows which do not have (close) neighbours:

# We compute the log and the spatial lag of some explanatory vriables:
flows_rem.red$log_send_electricity_use_pc <- log(flows_rem.red$send_Electricity_use_pc)
flows_rem.red$log_rec_electricity_use_pc <- log(flows_rem.red$rec_Electricity_use_pc)

# creating the spatial lagged variables  
flows_rem.red$lagged_log_rec_electricity_use_pc <- flows_rem.red$lagged_rec_electricity_use_pc <-  flows_rem.red$lagged_log_rec_pop <- 
flows_rem.red$lagged_rec_pop <- flows_rem.red$lagged_log_send_electricity_use_pc <- flows_rem.red$lagged_send_electricity_use_pc <- flows_rem.red$lagged_log_send_pop <- flows_rem.red$lagged_send_pop <- NA

flows_rem.red[!flows_rem.red$flow.no.neigh.od, c("lagged_log_rec_electricity_use_pc","lagged_rec_electricity_use_pc", "lagged_log_rec_pop", "lagged_rec_pop")] <-
  Wd.dgT_d%*%as.matrix(flows_rem.red[!flows_rem.red$flow.no.neigh.od, c("log_rec_electricity_use_pc","rec_Electricity_use_pc","log_rec_pop", "rec_population")] )

flows_rem.red[!flows_rem.red$flow.no.neigh.od, c("lagged_log_send_electricity_use_pc","lagged_send_electricity_use_pc","lagged_log_send_pop", "lagged_send_pop")] <-
  Wo.dgT_o%*%as.matrix(flows_rem.red[!flows_rem.red$flow.no.neigh.od, c("log_send_electricity_use_pc","send_Electricity_use_pc","log_sendpop", "send_population")] )

flows_rem.red <- flows_rem.red[!flows_rem.red$flow.no.neigh.od, ]

library("spdep")
w_rem_o_nb <- knn2nb(knearneigh(as(st_centroid(geom_send_ct), "Spatial"), k = 3))
w_rem_o <- listw2mat(nb2listw(w_rem_o_nb))

row.names(w_rem_o) <- row.names(geom_send_ct)
colnames(w_rem_o) <- row.names(geom_send_ct)

### Wd

w_rem_d_nb <- knn2nb(knearneigh(as(st_centroid(geom_rec_ct), "Spatial"), k = 3))
w_rem_d <- listw2mat(nb2listw(w_rem_d_nb))

# ADDED
row.names(w_rem_d) <- row.names(geom_rec_ct)
colnames(w_rem_d) <- row.names(geom_rec_ct)
# END OF ADDED

```


# Modeling

```{r}
# We need to load internal function for the estimation of the Spatial SDM flows models:

source("ftrace1.R")
source("sar_flow_2.R")
```


## Specification 1

```{r}
# prepare the variables we need
OX_remit <-  data.frame(unique(flows_rem.red[, c("origin", "log_send_electricity_use_pc", "log_sendpop", "send_island", 
"send_high_inf", "send_cont_mulFX", "send_FXsd", "send_WU_resid")]), row.names = "origin")
names(OX_remit) <- c("log_electricity_use_pc", "log_pop", "island", "high_inf", "cont_mulFX", "FXsd", "WU_resid")

DX_remit <- data.frame(unique(flows_rem.red[, c("dest", "log_rec_electricity_use_pc", "log_rec_pop", "rec_high_inf", "rec_polstab", "rec_landlocked", "rec_dum_nat_dis")]), row.names = "dest") 
names(DX_remit) <- c("log_electricity_use_pc", "log_pop", "high_inf", "polstab", "landlocked", "dum_nat_dis")

pairs_OD <- flows_rem.red[, c("contig", "comlang_off", "comcur", "colony", "logg")]
y_od <- flows_rem.red[, c("origin", "dest", "logy")]

W_o_remit <- Wo.dgT_o
W_d_remit <- Wd.dgT_d
save(OX_remit, DX_remit, pairs_OD, y_od, W_o_remit, W_d_remit, file = "my_remit.RData")
```


### Ordinary Least Squares

```{r,echo=FALSE}
# + send_med_inf + rec_med_inf

formula_lm_sdm <- as.formula("logy ~  log_send_electricity_use_pc + log_sendpop + 
   send_island + send_high_inf  + send_cont_mulFX + send_FXsd + 
   send_WU_resid + lagged_log_send_electricity_use_pc + 
   log_rec_electricity_use_pc + log_rec_pop + 
   rec_high_inf  + rec_polstab + rec_landlocked + rec_dum_nat_dis + lagged_log_rec_electricity_use_pc +
   contig + 
   comlang_off + comcur + colony + logg")


fito <- lm(formula_lm_sdm, data = flows_rem.red)
S <- summary(fito)
# We print the results:
# S
# We store the results for exporting later to LaTeX:
lm_res <- rbind(c("", ""), c("", ""), round(cbind(S$coefficients[1:20 ,1], 
                                                  S[["coefficients"]][1:20, "t value"]), digits = 3),
              round(cbind(S$coefficients[21, 1], S[["coefficients"]][21, "t value"]), digits = 3))
row.names(lm_res) <- NULL
```

## Bayesian model


### With $W_d$

```{r, eval = T,echo=FALSE,include=FALSE}
# Formula SDM
system.time(sdm_bay_Wd_2017_13 <- 
              sar_flow_2(x = flows_rem.red[, 
               c("log_send_electricity_use_pc", "log_sendpop", "send_island",
                 "send_high_inf",   
                 "send_cont_mulFX", "send_FXsd",  "send_WU_resid","lagged_log_send_electricity_use_pc", 
                 "log_rec_electricity_use_pc", "log_rec_pop",
                 "rec_high_inf", "rec_polstab","rec_landlocked",
                 "rec_dum_nat_dis", "lagged_log_rec_electricity_use_pc",
                 "contig", "comlang_off", "comcur", "colony")], 
    y = flows_rem.red$logy, 
    g = flows_rem.red$logg, 
    W_d = Wd.dgT_d,  
    model = "model_2"))


### With $W_o$
system.time(sdm_bay_Wo_2017_13 <- 
              sar_flow_2(x = flows_rem.red[ , 
             c("log_send_electricity_use_pc", "log_sendpop", "send_island",
                 "send_high_inf",   
                 "send_cont_mulFX", "send_FXsd",  "send_WU_resid","lagged_log_send_electricity_use_pc", 
                 "log_rec_electricity_use_pc", "log_rec_pop",
                 "rec_high_inf","rec_polstab","rec_landlocked",
                 "rec_dum_nat_dis", "lagged_log_rec_electricity_use_pc",
                 "contig", "comlang_off", "comcur", "colony")],
  y = flows_rem.red$logy, 
  g = flows_rem.red$logg, 
  W_o = Wo.dgT_o,  
  model = "model_3"))

 
### With $W_o$ and $W_d$

system.time(sdm_bay_Wod_2017_13 <- sar_flow_2(
  x = flows_rem.red[!flows_rem.red$flow.no.neigh.od, 
           c("log_send_electricity_use_pc", "log_sendpop", "send_island",
                 "send_high_inf",   
                 "send_cont_mulFX", "send_FXsd",  "send_WU_resid","lagged_log_send_electricity_use_pc", 
                 "log_rec_electricity_use_pc", "log_rec_pop",
                 "rec_high_inf","rec_polstab","rec_landlocked",
                 "rec_dum_nat_dis", "lagged_log_rec_electricity_use_pc",
                 "contig", "comlang_off", "comcur", "colony")],
  y = flows_rem.red$logy[!flows_rem.red$flow.no.neigh.od], 
  g = flows_rem.red$logg[!flows_rem.red$flow.no.neigh.od], 
  W_o = Wo.dgT_o, W_d = Wd.dgT_d,  
  model = "model_7"))

# We print the results:

sdm_bay_Wd_2017_b_sum <- summary_bayesian_sar_flow(sdm_bay_Wd_2017_13)
ndraw <- length(sdm_bay_Wd_2017_13$psave)
sdm_bay_Wo_2017_b_sum <- summary_bayesian_sar_flow(sdm_bay_Wo_2017_13)
sdm_bay_Wod_2017_b_sum <- summary_bayesian_sar_flow(sdm_bay_Wod_2017_13)

N <- nrow(flows_rem.red[!flows_rem.red$flow.no.neigh.od, ])
require("ggridges")

# Table short


```

## Print the results: complete results

```{r,echo=FALSE}

# var_names <- c("Spat par O", "Spat par D", "Intercept", "Log Electric Use PC O", 
#                  "Log Pop O", "Island O", "High Inf O",  "Multiple FX O", 
#                  "FX Vol O", "Transfer Cost O", "Lagged Log Electric Use PC O",
#                 "Log Electric Use PC D", "Log Pop D", 
#                  "High Inf D",  "Pol Stab D", "Landlocked D", "Natural Disaster D",
#                "Lagged Log Electric Use PC D",
#                "Com Border", "Com Language", "Com Currency", "Colony", "Distance")

var_names <- c("Spatial parameter O", "Spatial parameter D", "Intercept", "Log(Electric Use PC O)", 
                 "Log(Population O)", "Island O", "High Inflation O",  "Multiple Foreign Exchanges O", 
                 "Foreign Exchange Volatility O", "Transfer Cost O", "Lagged Log(Electric Use PC O)",
                "Log(Electric Use PC D)", "Log(Population D)", 
                 "High Inflation D",  "Political Stability D", "Landlocked D", "Natural Disaster D",
               "Lagged Log(Electric Use PC D)",
               "Common Border", "Common Language", "Common Currency", "Colony", "Distance")
sdm_bay_Wo_2017_b_sum <- rbind(sdm_bay_Wo_2017_b_sum[1, ],
                             c("", "", "", ""), sdm_bay_Wo_2017_b_sum[-1, ])
sdm_bay_Wd_2017_b_sum <- rbind(c("", "", "", ""), sdm_bay_Wd_2017_b_sum)
nrow(sdm_bay_Wd_2017_b_sum)


result_fin_2017_13 <- data.frame(cbind(var_names, lm_res[, 1], lm_res[, 2],
                                    as.numeric(round(as.numeric(sdm_bay_Wo_2017_b_sum[, 1]), 
                                                     digits = 3)),
                                    as.numeric(round(as.numeric(sdm_bay_Wo_2017_b_sum[, 4]), 
                                                     digits = 3)),
                                    as.numeric(round(as.numeric(sdm_bay_Wd_2017_b_sum[, 1]), 
                                                     digits = 3)),
                                    as.numeric(round(as.numeric(sdm_bay_Wd_2017_b_sum[, 4]), 
                                                     digits = 3)),
                                    as.numeric(round(as.numeric(sdm_bay_Wod_2017_b_sum[, 1]), 
                                                     digits = 3)),
                                    as.numeric(round(as.numeric(sdm_bay_Wod_2017_b_sum[, 4]), 
                                                     digits = 3))))
names(result_fin_2017_13) <- c("Variables", "OLM", "T-stat", "Est Wo", "T-stat", "Est Wd", "T-stat",
                               "Wo Wd", "T-stat")
rownames(result_fin_2017_13)<-NULL

knitr::kable(result_fin_2017_13)

```

## Short version to be presented in the paper

```{r,echo=FALSE}

## Print the results
result_fin_2017_short <- data.frame(cbind(var_names, lm_res[, 1], lm_res[, 2],
                                    as.numeric(round(as.numeric(sdm_bay_Wod_2017_b_sum[, 1]), 
                                                     digits = 3)),
                                    as.numeric(round(as.numeric(sdm_bay_Wod_2017_b_sum[, 4]), 
                                                     digits = 3))))

names(result_fin_2017_short) <- c("Variables", "OLM", "T-stat", 
                               "Wo Wd", "T-stat")
rownames(result_fin_2017_short)<-NULL

# Table short
knitr::kable(result_fin_2017_short)
```


